---
title: "Kratom and Violent Crime"
author: "J.D. Dato"
format: html
editor: visual
execute: 
  echo: false
toc: True
theme: solar
smooth-scroll: true
---

# Kratom and Crime

***Consider that this data set is an example and not reflective of any real relationship that could exist between Kratom and crime***

## 3-D Scatterplot

 This 3D Scatterplot set shows a possible relationship between violent crimes and Kratom sales in/around a local head shop. The colors show how many transactions were made in one day. The Z-axis (up-down), shows the level of violent crime. The Y and X axis show the Units of Kratom sold, and the dollar amount of total sales, respectively. ***3D Visualizations are a great way to show 3+ variables all at once, in a way that is still simple to understand.***

```{r, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE, results='hide'}
library(tidyverse)
library(ggplot2)
library(plyr)
library(plotly)
library(flexdashboard)
library(readr)
library(lattice)
library(reshape2)
KratomSales1 <- read.csv("KratomSales1.csv", header = TRUE)
litchart <- ggplot(data = KratomSales1, aes(x=TotalSales, y= ViolentCrime, color=Units)) +
  geom_point() + geom_smooth(method = lm, fill='green', se=TRUE, alpha=0.09) + labs(title='Kratom Sales Compared to Violent Crime', 
            subtitle = "Compared to both units sold and total sales amount",
            y= 'Violent Crime Rate for the Week',
            x='Total Kratom Sales for the Week') 


```

```{r, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
plot_ly(KratomSales1, x = ~TotalSales, y = ~Units, z = ~ViolentCrime, color = ~Transactions,
        colors = c('#FFC000', '#228B22'))
```

## Correlation Matrix

This is a correlation matrix. Each square represents the pearson correlation r value between the two intersecting variables. The strength/direction of the correlation can be determined from viewing the legend on the right side and comparing it to the square representing the variables in question.

```{r, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
KratomSales2 <- as.data.frame(KratomSales1)
KratomSales3 <- KratomSales2[-c(1,2,7)]
corr_mat <- round(cor(KratomSales3),2)

melted_corr_mat <- melt(corr_mat)
gx <- ggplot(data = melted_corr_mat, aes(x=Var1, y=Var2, fill = value))+
  geom_tile() +  scale_fill_gradient(low = "#228B22",
                                     high = "#FFC000",
                                     guide = "colorbar")
ggplotly(gx)
```
